Zero-shot task adaptation by homoiconic meta-mapping

### 2019-09-07 writing/embedded_meta_learning_appendix.tex
Fixing some references
length cuts

\subsection{Polynomials}
\begin{table}[H]
\scriptsize
\centering
\begin{tabular}{@{\extracolsep{5pt}} ccccc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
named\_run\_type & is\_new & mean\_loss & boot\_CI\_low & boot\_CI\_high \\
\hline \\[-1.8ex]
EML & Trained & 0.015 & 0.012 & 0.018 \\
EML & Held out & 0.246 & 0.188 & 0.308 \\
Untrained EML network & Trained & 5.735 & 4.823 & 6.74 \\
Untrained EML network & Held out & 5.968 & 4.984 & 6.991 \\
\hline \\[-1.8ex]
\end{tabular}
\caption{Table for basic meta-learning, figure \ref{poly_basic_results}}
\end{table}

\begin{table}[H]
\scriptsize
\centering
\begin{tabular}{@{\extracolsep{5pt}} ccccc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
named\_run\_type & result\_type & mean\_loss & boot\_CI\_low & boot\_CI\_high \\
\hline \\[-1.8ex]
EML & Trained mapping, on trained task & 0.094 & 0.091 & 0.098 \\
EML & Trained mapping, on held-out task & 1.721 & 1.419 & 2.115 \\
EML & Held-out mapping, on trained task & 1.28 & 1.213 & 1.35 \\
EML & Held-out mapping, on held-out task & 1.775 & 1.706 & 1.846 \\
Untrained EML network & Trained mapping, on trained task & 12.998 & 11.689 & 14.381 \\
Untrained EML network & Trained mapping, on held-out task & 15.002 & 13.39 & 16.83 \\
Untrained EML network & Held-out mapping, on trained task & 8.36 & 7.898 & 8.833 \\
Untrained EML network & Held-out mapping, on held-out task & 8.786 & 8.317 & 9.27 \\
\hline \\[-1.8ex]
\end{tabular}
\caption{Table for meta-mapping results from examples, figure \ref{poly_meta_map_results_examples}}
\end{table}

\begin{table}[H]
\scriptsize
\centering
\begin{tabular}{@{\extracolsep{5pt}} ccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
named\_run\_type & result\_type & mean\_loss & boot\_CI\_low & boot\_CI\_high \\ 
\hline \\[-1.8ex] 
Language & Trained mapping, on trained task & 0.515 & 0.483 & 0.552 \\ 
Language & Trained mapping, on held-out task & 2.244 & 1.921 & 2.623 \\ 
Language & Held-out mapping, on trained task & 2.072 & 1.958 & 2.19 \\ 
Language & Held-out mapping, on held-out task & 2.35 & 2.254 & 2.447 \\ 
Untrained EML network & Trained mapping, on trained task & 13.328 & 11.977 & 14.823 \\ 
Untrained EML network & Trained mapping, on held-out task & 15.313 & 13.602 & 17.354 \\ 
Untrained EML network & Held-out mapping, on trained task & 8.205 & 7.795 & 8.662 \\ 
Untrained EML network & Held-out mapping, on held-out task & 8.625 & 8.131 & 9.104 \\ 
\hline \\[-1.8ex] 
\end{tabular}
\caption{Table for meta-mapping results from language, figure \ref{poly_meta_map_results_language}}
\end{table}

@@

\begin{table}[H]
\scriptsize
\centering
\begin{tabular}{@{\extracolsep{5pt}} cccc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
named\_run\_type & named\_meta\_task & is\_new & expected\_reward \\
\hline \\[-1.8ex]
Targeted 10\% holdout & Switch suits & Trained & 0.298 \\
Targeted 10\% holdout & Switch suits & Held out & 0.368 \\
Targeted 10\% holdout & Losers & Trained & -0.446 \\
Targeted 10\% holdout & Losers & Held out & -0.463 \\
Random 50\% holdout & Switch suits & Trained & 0.37 \\
Random 50\% holdout & Switch suits & Held out & 0.278 \\
Random 50\% holdout & Losers & Trained & -0.465 \\
Random 50\% holdout & Losers & Held out & -0.444 \\
\hline \\[-1.8ex]
\end{tabular}
\caption{Table of rewards if system ignored meta-mapping, colored dashed lines in figure \ref{cards_meta_map_results}}
\end{table}

>>>>

\subsection{Polynomials}
\begin{table}[H]
\scriptsize
\centering
\begin{tabular}{@{\extracolsep{5pt}} ccccc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
named\_run\_type & is\_new & mean\_loss & boot\_CI\_low & boot\_CI\_high \\
\hline \\[-1.8ex]
HoMM & Trained & 0.015 & 0.012 & 0.018 \\
HoMM & Held out & 0.246 & 0.188 & 0.308 \\
Untrained HoMM network & Trained & 5.735 & 4.823 & 6.74 \\
Untrained HoMM network & Held out & 5.968 & 4.984 & 6.991 \\
\hline \\[-1.8ex]
\end{tabular}
\caption{Table for basic meta-learning, figure \ref{poly_basic_results}}
\end{table}

\begin{table}[H]
\scriptsize
\centering
\begin{tabular}{@{\extracolsep{5pt}} ccccc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
named\_run\_type & result\_type & mean\_loss & boot\_CI\_low & boot\_CI\_high \\
\hline \\[-1.8ex]
HoMM & Trained mapping, on trained task & 0.094 & 0.091 & 0.098 \\
HoMM & Trained mapping, on held-out task & 1.721 & 1.419 & 2.115 \\
HoMM & Held-out mapping, on trained task & 1.28 & 1.213 & 1.35 \\
HoMM & Held-out mapping, on held-out task & 1.775 & 1.706 & 1.846 \\
Untrained HoMM network & Trained mapping, on trained task & 12.998 & 11.689 & 14.381 \\
Untrained HoMM network & Trained mapping, on held-out task & 15.002 & 13.39 & 16.83 \\
Untrained HoMM network & Held-out mapping, on trained task & 8.36 & 7.898 & 8.833 \\
Untrained HoMM network & Held-out mapping, on held-out task & 8.786 & 8.317 & 9.27 \\
\hline \\[-1.8ex]
\end{tabular}
\caption{Table for meta-mapping results from examples, figure \ref{poly_meta_map_results_examples}}
\end{table}

\begin{table}[H]
\scriptsize
\centering
\begin{tabular}{@{\extracolsep{5pt}} ccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
named\_run\_type & result\_type & mean\_loss & boot\_CI\_low & boot\_CI\_high \\ 
\hline \\[-1.8ex] 
Language & Trained mapping, on trained task & 0.515 & 0.483 & 0.552 \\ 
Language & Trained mapping, on held-out task & 2.244 & 1.921 & 2.623 \\ 
Language & Held-out mapping, on trained task & 2.072 & 1.958 & 2.19 \\ 
Language & Held-out mapping, on held-out task & 2.35 & 2.254 & 2.447 \\ 
Untrained HoMM network & Trained mapping, on trained task & 13.328 & 11.977 & 14.823 \\ 
Untrained HoMM network & Trained mapping, on held-out task & 15.313 & 13.602 & 17.354 \\ 
Untrained HoMM network & Held-out mapping, on trained task & 8.205 & 7.795 & 8.662 \\ 
Untrained HoMM network & Held-out mapping, on held-out task & 8.625 & 8.131 & 9.104 \\ 
\hline \\[-1.8ex] 
\end{tabular}
\caption{Table for meta-mapping results from language, figure \ref{poly_meta_map_results_language}}
\end{table}

@@

\begin{table}[H]
\scriptsize
\centering
\begin{tabular}{@{\extracolsep{5pt}} cccc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
named\_run\_type & named\_meta\_task & is\_new & expected\_reward \\
\hline \\[-1.8ex]
Targeted 10\% holdout & Switch suits & Trained & 0.298 \\
Targeted 10\% holdout & Switch suits & Held out & 0.368 \\
Targeted 10\% holdout & Losers & Trained & -0.446 \\
Targeted 10\% holdout & Losers & Held out & -0.463 \\
Random 50\% holdout & Switch suits & Trained & 0.37 \\
Random 50\% holdout & Switch suits & Held out & 0.278 \\
Random 50\% holdout & Losers & Trained & -0.465 \\
Random 50\% holdout & Losers & Held out & -0.444 \\
\hline \\[-1.8ex]
\end{tabular}
\caption{Table of rewards if system ignored meta-mapping, colored dashed lines in figure \ref{cards_meta_map_results_examples}}
\end{table}

### 2019-09-07 writing/embedded_meta_learning_main.tex
length cuts

\vspace{-0.25em}
\section{Meta-mapping}
\vspace{-0.5em} % section
%Can cut "the task"s in go sentence if needed
We propose the meta-mapping challenge. We define a meta-mapping as a task, or mapping, that takes a task as an input, output, or both. These include mapping from tasks to language (explaining), mapping from language to tasks (following instructions), and mapping from tasks to tasks (adapting behavior). While the first two categories have been partially addressed in prior work \citep[e.g.][]{Hermann2017, Co-Reyes2019}, the latter is more novel. (We discuss the relationship between our work and prior work in section \ref{sec_discussion}.) \par
We argue that task-to-task meta-mappings are a useful way to think about human-like flexibility, because a great deal of our rapid adaptation is from a task to some variation on that task. For example, the task of playing go on a large board is closely related to the task of playing go on a small board. Humans can exploit this to immediately play well on a different board, but deep learning models generally have no way to achieve this. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to lose, or trying to achieve some orthogonal goal. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not fully capture human flexibility. Given appropriate conditioning (see below), our architecture can adapt to substantial task alterations zero-shot, that is, without seeing a single example from the new task \citep{Lake2016}. We suggest that meta-mappings offer a way to understand this flexibility, and a way for deep-learning models to achieve it.\par
This adaptation can be cued in different ways, for example by examples of similar task mappings (after winning and losing at poker, try to lose at blackjack) or simply by natural-language instructions (``try to lose at blackjack'', see appendix \ref{app_why_meta_mapping} for a comparison to conditioning on a task description.). Achieving the flexibility to adapt to meta-map to new tasks will be an important step toward more general intelligence -- intelligence that is not limited to exactly the training domains it has seen. \par

@@

More formally, we treat functions and data as entities of the same type. From this perspective, the data points that one function receives can themselves be functions\footnote{Indeed, any data point can be represented as a constant function that outputs the data point.}. The key insight is that then our architecture can transform data points\footnote{Where ``data'' is a quite flexible term. The approach is agnostic to whether the learning is supervised or reinforcement learning, whether inputs are images or natural language, etc.} to perform basic tasks (as is standard in machine learning), but it can also transform these task functions to adapt to new tasks. This is related to the idea of homoiconicity, defined above, and the concept of higher-order functions. Under this perspective, basic tasks and meta-mappings from task to task are really the same type of problem. The functions at one level of abstraction (the basic tasks) become inputs and outputs for higher-level functions at the next level of abstraction (meta-mapping between tasks). \par

@@

\paragraph{Basic meta-learning:} To perform a basic task, input and target encoders ($\mathcal{I}$ and $\mathcal{T}$) are used to embed individual pairs from an example dataset \(D_1\), to form a dataset of example (input, output) tuples. These examples are fed to $\mathcal{M}$, which produces a function embedding (via a deep neural network including an element-wise max across the examples). This function embedding is mapped through they hyper network $\mathcal{H}$ to parameterize $F$, and then $F$ is used to process a dataset of embedded probe inputs, and $\mathcal{O}$ to map the resultant embeddings to outputs. This system can be trained end-to-end on targets for the second dataset. Having two distinct datasets forces generalization at the meta-learning level, see appendix \ref{app_clarifying_holdouts}. See appendix \ref{app_model_details} for detailed architecture, operation, and hyper-parameters. \par 

@@

Language is fundamental to human flexibility, and we would be remiss to ignore it. In many of the intuitive illustrations we've given, the examples of the meta-mapping are implicit in your prior knowledge about the world. For example, ``try to lose at go'' does not give explicit examples of the ``lose'' meta-mapping, but rather relies on prior knowledge of what losing means. This is a much more efficient way to achieve zero-shot performance on a new losing task. In order to replicate this, we trained the HoMM system with both meta-mappings based on examples, and meta-mappings based on language. In the language-based meta-mappings, a language input identifying the meta-mapping (but not the basic task to apply it to) is encoded by a language encoder, and then provided as the input to $\mathcal{H}$ (instead of an output from $\mathcal{M}$). The meta-mapping then proceeds as normal --- $\mathcal{H}$ parameterizes $F$, which is used to transform the embedding of the input task to produce an embedding for the target. \par

@@

We've highlighted a new type of flexibility in the form of meta-mapping -- mapping between tasks. Meta-mapping can produce zero-shot performance on novel tasks, based on their relationship to old tasks. However, achieving this type of flexibility is difficult in architectures that separate computation from the knowledge of what is being computed. To address this, our HoMM approach explicitly represents tasks by function embeddings, and derives the computation from these embeddings via a HyperNetwork. This allows for our key insight: by embedding tasks and individual data points in a shared latent space, the same meta-learning architecture can be used both to transform data for basic tasks, and to transformat task embeddings to adapt to task variations. This approach is parsiomious, because it uses the same networks for basic- and meta-mappings. Perhaps because of this, the requisite meta-task sample complexity is small; we showed generalization to unseen meta-mappings after training on only 20 meta-mappings. That is, we are able to achieve zero-shot performance on a novel task based on a relationship between tasks that was never trained on. This is a step closer to human-level flexibility.\par

>>>>

\vspace{-0.25em}
\section{Meta-mapping}
\vspace{-0.5em} % section
%Can cut "the task"s in go sentence if needed
We propose the meta-mapping challenge. We define a meta-mapping as a task, or mapping, that takes a task as an input, output, or both. These include mapping from tasks to language (explaining), mapping from language to tasks (following instructions), and mapping from tasks to tasks (adapting behavior). While the first two categories have been partially addressed in prior work \citep[e.g.][]{Hermann2017, Co-Reyes2019}, the latter is more novel. (We discuss the relationship between our work and prior work in section \ref{sec_discussion}.) \par
We argue that task-to-task meta-mappings are a useful way to think about human-like flexibility, because a great deal of our rapid adaptation is from a task to some variation on that task. For example, the task of playing go on a large board is closely related to the task of playing go on a small board. Humans can exploit this to immediately play well on a different board, but deep learning models generally have no way to achieve this. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to lose, or trying to achieve some orthogonal goal. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not fully capture human flexibility. Given appropriate conditioning (see below), our architecture can adapt to substantial task alterations zero-shot, that is, without seeing a single example from the new task \citep{Lake2016}. We suggest that meta-mappings offer a way to understand this flexibility, and a way for deep-learning models to achieve it.\par
This adaptation can be cued in different ways, for example by examples of similar task mappings (after winning and losing at poker, try to lose at blackjack) or simply by natural-language instructions (``try to lose at blackjack''). Achieving the flexibility to adapt to meta-map to new tasks will be an important step toward more general intelligence -- intelligence that is not limited to exactly the training domains it has seen. \par

@@

More formally, we treat functions and data as entities of the same type. From this perspective, the data points that one function receives can themselves be functions\footnote{Indeed, any data point can be represented as a constant function that outputs the data point.}. The key insight is that then our architecture can transform data points\footnote{Where ``data'' is a quite flexible term. The approach is agnostic to whether the learning is supervised or reinforcement learning, whether inputs are images or natural language, etc.} to perform basic tasks (as is standard in machine learning), but it can also transform these task functions to adapt to new tasks. This is related to the concepts of homoiconicity, defined above, and higher-order functions. Under this perspective, basic tasks and meta-mappings from task to task are really the same type of problem. The functions at one level of abstraction (the basic tasks) become inputs and outputs for higher-level functions at the next level of abstraction (meta-mapping between tasks). \par

@@

\paragraph{Basic meta-learning:} To perform a basic task, input and target encoders ($\mathcal{I}$ and $\mathcal{T}$) are used to embed individual pairs from an example dataset \(D_1\), to form a dataset of example (input, output) tuples. These examples are fed to $\mathcal{M}$, which produces a function embedding (via a deep neural network, including an element-wise max across the examples). This function embedding is mapped through they hyper network $\mathcal{H}$ to parameterize $F$, and then $F$ is used to process a dataset of embedded probe inputs, and $\mathcal{O}$ to map the resultant embeddings to outputs. This system can be trained end-to-end on target outputs for the probes. Having two distinct datasets forces generalization at the meta-learning level, see appendix \ref{app_clarifying_holdouts}. See appendix \ref{app_model_details} for detailed architecture, and hyper-parameters. \par 

@@

Language is fundamental to human flexibility. Often the examples of the meta-mapping are implicit in prior knowledge about the world that is cued by language. For example, ``try to lose at go'' does not give explicit examples of the ``lose'' meta-mapping, but rather relies on prior knowledge of what losing means. This is a much more efficient way to cue a known meta-mapping. In order to replicate this, we trained the HoMM system with both meta-mappings based on examples, and meta-mappings based on language. In the language-based meta-mappings, a language input identifying the meta-mapping (but not the basic task to apply it to) is encoded by a language encoder, and then provided as the input to $\mathcal{H}$ (instead of an output from $\mathcal{M}$). The meta-mapping then proceeds as normal --- $\mathcal{H}$ parameterizes $F$, which is used to transform the embedding of the input task to produce an embedding for the target. \par

@@

We've highlighted a new type of flexibility in the form of meta-mapping between tasks. Meta-mapping can produce zero-shot performance on novel tasks, based on their relationship to old tasks. However, achieving this flexibility is difficult in architectures that separate computation from the knowledge of what is being computed. To address this, our Homoiconic Meta-Mapping (HoMM) approach explicitly represents tasks by function embeddings, and derives the computation from these embeddings via a HyperNetwork. This allows for our key insight: by embedding tasks and individual data points in a shared latent space, the same meta-learning architecture can be used both to transform data for basic tasks, and to transformat task embeddings to adapt to task variations. This approach is parsiomious, because it uses the same networks for basic- and meta-mappings. Perhaps because of this, the requisite meta-task sample complexity is small; we showed generalization to unseen meta-mappings after training on only 20 meta-mappings. That is, we are able to achieve zero-shot performance on a novel task based on a held-out relationship between tasks. This is a step closer to human-level flexibility.\par
